---
title: "Kaplan–Meier Survival Analysis"
author: "Sebastian Caro"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

# Overview

This notebook demonstrates the workflow used to perform Kaplan–Meier survival analysis, compare survival curves, and estimate median survival.

# Load required packages

```{r setup, message=FALSE, warning=FALSE}
library(tidyr)
library(flextable)
library(officer)
library(openxlsx)
library(dplyr)

library(survival)
library(survminer)
library(survMisc)

library(viridis)
library(ggsci)
```

# Kaplan–Meier estimator

The Kaplan–Meier estimator is fitted using the `survfit()` function. 

> **Note:** Replace `time` and `status` with your actual time-to-event and event indicator column names.

```{r fit-km}
# Fitting the survival curve based on genetic fusion status
fit_km <- survfit(
  Surv(time, status) ~ gen_fusion,
  data = my_data
)
```

# Life table

Display the life table generated by the Kaplan–Meier model to inspect survival probabilities at specific time points.

```{r summary-km}
summary(fit_km)
```

# Kaplan–Meier Plot

Visualizing the survival curves using `survminer`. This plot includes the automatic log-rank p-value, a risk table, and a scientific color palette.

```{r plot-km, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
ggsurvplot(
  fit_km,
  data = my_data,
  pval = TRUE,             # Adds the log-rank test p-value automatically
  risktable = TRUE,        # Adds the number-at-risk table below the plot
  palette = "npg",         # Nature Publishing Group color palette
  ggtheme = theme_minimal(),
  legend.labs = levels(as.factor(my_data$gen_fusion)) # Dynamic legends based on your groups
)
```

# Tarone–Ware test

The Tarone–Ware test provides an alternative comparison between survival curves by assigning greater weight to earlier events, which is useful if early divergence is suspected.

```{r tarone, message=FALSE, warning=FALSE}
fit_ten <- ten(fit_km)
comp(fit_ten)
```

# Log-rank test

The Log-rank test evaluates whether survival curves differ significantly between groups, giving equal weight to all events across the timeline. Here we evaluate survival by aneuploidy status.

```{r logrank}
log_rank <- survdiff(
  Surv(time, status) ~ aneuplo,
  data = my_data
)
log_rank
```

# Calculate p-value

The p-value is extracted from the chi-square statistic of the log-rank test using a numerically stable function for high precision.

```{r pvalue}
chisq_stat <- log_rank$chisq
df <- length(log_rank$n) - 1

# Numerically stable method for extremely small p-values
p_val <- pchisq(chisq_stat, df, lower.tail = FALSE)

cat("Log-Rank Test Result\n")
cat("--------------------\n")
cat("Chi-square         :", round(chisq_stat, 2), "\n")
cat("Degrees of freedom :", df, "\n")
cat("P-value            :", format.pval(p_val, digits = 4), "\n")
```

# Median survival

Estimate the median survival time (the time point at which the survival probability drops to 50%) along with its 95% confidence interval.

```{r median}
median_surv <- surv_median(fit_km)
median_surv
```

# Interpretation

- **Kaplan–Meier Estimator:** Provides the descriptive steps of survival probability over the follow-up period.
- **Log-rank Test:** Standard non-parametric test to check for statistical differences between the curves.
- **Tarone–Ware Test:** More sensitive to early clinical events or early differences during follow-up.
- **Median Survival:** If reported as `NA`, it implies that the survival curve did not drop below 0.5 (fewer than 50% of the patients in that group experienced the event during the study period).
